---
title: "FibroblastDEgenes"
author: "A.DeMartin"
date: "2025-08-18"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#### load packages
```{r load packages}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(here)
```

#### load merged object allFb
```{r}
## load object merged allFb naive and d4
basedir <- here()
fileNam <- paste0(basedir, "/data/merged_allFb_naiveANDd4_seurat.rds")
seuratFb <- readRDS(fileNam)
```

#### subset wt
```{r}
table(seuratFb$cond2)
seuratwt <- subset(seuratFb, cond2 == "WT")
table(seuratwt$cond2)

seuratFb <- seuratwt
```

#### exclude glial cells
```{r}
## filter glial cells
table(seuratFb$clustername)
seuratFb_noglial <- subset(seuratFb, clustername == "Glial", invert = TRUE)
table(seuratFb_noglial$clustername)

seuratFb <- seuratFb_noglial
```

#### set color vectors 
```{r}
colcond2 <- c("#202547","#BE3144")
names(colcond2) <- c("WT", "cko")

colcond3 <- c("#B45B5C","#628395","#BF782D")
names(colcond3) <- c("infectedD4", "naive", "granulomaD4")

colclustername <- c("#355C7D", "#8E9B97","#99B898","#E84A5F","#0F1F38","#F8B195","#727077","#FF847C","#C06C84")
names(colclustername) <- c("PdgfraloFb1" ,"PdgfraloFb2", "Trophocytes","PdgfraloFb3", "Telocytes","Myocytes", "Pdgfrahi", "PdgfraloFb4", "Thy1Fb")
```

#### counts
```{r}
table(seuratFb_noglial$orig.ident)
table(seuratFb_noglial$cond3)
```

#### calculate DE genes according to cond3
```{r, include=TRUE, eval=FALSE}
## calculate DE genes cond3
Idents(seuratFb) <- seuratFb$cond3
levels(seuratFb)

DEGenesFbcond3 <- FindAllMarkers(seuratFb, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01) 

write.table(DEGenesFbcond3, 
            file= paste0(basedir,"/data/DEGenesFbcond3.txt"),
            sep="\t",
            quote=F,
            row.names=F,
            col.names=T)
```

### GSEA Fb Gran
#### dotplot
```{r GSEA Fb gran, fig.height=12, fig.width=12}
## load DEGenesFbcond3
DEGenesFbcond3 <- read.delim(paste0(basedir,"/data/DEGenesFbcond3.txt"), header = TRUE, sep = "\t")

## genes upregulated in Gran
DEGenesGran <- dplyr::filter(DEGenesFbcond3, cluster == "granulomaD4")
DEGenesGran <- DEGenesGran %>%
  mutate(Gene=gsub("^.*\\.", "", gene))  %>%
  mutate(EnsID=gsub("\\..*","", gene))

## GSEA Gran
egoGran <- enrichGO(gene = unique(DEGenesGran$EnsID),
                          OrgDb = org.Mm.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)
egoGran <- setReadable(egoGran, OrgDb = org.Mm.eg.db)
dotplot(egoGran, showCategory=30)
```

#### barplot sel pathways Gran
```{r barplot GSEA Fb gran, fig.height=5, fig.width=12}
## select pahtways and filter file
selGran <- c("leukocyte migration","regulation of innate immune response", "mitochondrial translation", "response to interferon-beta")
egoGran_fil <- egoGran%>% filter(egoGran@result$Description %in% selGran)
egoGran_fil_2 <- pairwise_termsim(egoGran_fil)

## make ggbarplot
p <- ggbarplot(egoGran_fil_2@result, x = "Description", y = "Count", fill = "#BF782D", color= "#BF782D", sort.val = "asc", orientation = "horizontal")
p
```

### signature umaps Gran
```{r}
## convert seurat object to sce object
Idents(seuratFb) <- seuratFb$cond3
sce <- as.SingleCellExperiment(seuratFb)
genes <- data.frame(geneID=rownames(sce)) %>% mutate(gene=gsub(".*\\.", "", geneID))
pal = colorRampPalette(c("#053061", "#2166ac", "#f7f7f7", "#f4a582", "#b2183c", "#85122d"))

## subset naive and convert to sce
seuratFbnaive <- subset(seuratFb, cond3 == "naive")
DimPlot(seuratFbnaive, reduction= "umap", cols = colcond3) 
sceFbnaive <- as.SingleCellExperiment(seuratFbnaive)

## subset SI and convert to sce
seuratFbinf <- subset(seuratFb, cond3 == "infectedD4")
DimPlot(seuratFbinf, reduction= "umap", cols = colcond3) 
sceFbinf <- as.SingleCellExperiment(seuratFbinf)

## subset Gran and convert to sce
seuratFbgran <- subset(seuratFb, cond3 == "granulomaD4")
DimPlot(seuratFbgran, reduction= "umap", cols = colcond3) 
sceFbgran <- as.SingleCellExperiment(seuratFbgran)
```

#### leukocyte migration
```{r plot GSEA leukocyte migration}
ego1 <- dplyr::filter(egoGran@result,egoGran@result$Description=="leukocyte migration")
g1 <- ego1$geneID
Str <-(g1)
StrSub <- strsplit(Str, "/")
df <- as.data.frame(StrSub)
colnames(df) <- c("gene")

signGenes <- genes %>% dplyr::filter(gene %in% df$gene)
## make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.6))
sceSub$sign2[which(sceSub$sign > 0.6)] <- 0.6
## check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.6))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 0.6)] <- 0.6
## check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.6))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 0.6)] <- 0.6
## check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.6))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 0.6)] <- 0.6
## check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

#### regulation of innate immune response
```{r plot GSEA positive regualtion of innate immune response}
ego1 <- dplyr::filter(egoGran@result,egoGran@result$Description=="regulation of innate immune response")
g1 <- ego1$geneID
Str <-(g1)
StrSub <- strsplit(Str, "/")
df <- as.data.frame(StrSub)
colnames(df) <- c("gene")

## make a count matrix of signature genes
signGenes <- genes %>% dplyr::filter(gene %in% df$gene)

sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.7))
sceSub$sign2[which(sceSub$sign > 0.7)] <- 0.7
## check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2") + sc +
  theme(legend.position = "none", point_size = 1)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.7))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 0.7)] <- 0.7
## check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.7))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 0.7)] <- 0.7
## check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.7))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 0.7)] <- 0.7
## check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

### GSEA Fb SI
#### dotplot
```{r GSEA Fb SI, fig.height=12, fig.width=12}
## genes upregulated in SI
DEGenesinf <- dplyr::filter(DEGenesFbcond3, cluster == "infectedD4")
DEGenesinf <- DEGenesinf %>%
  mutate(Gene=gsub("^.*\\.", "", gene))  %>%
  mutate(EnsID=gsub("\\..*","", gene))

## GSEA SI
egoinf <- enrichGO(gene = unique(DEGenesinf$EnsID),
                          OrgDb = org.Mm.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)
egoinf <- setReadable(egoinf, OrgDb = org.Mm.eg.db)
dotplot(egoinf, showCategory=30)
```

#### barplot sel pathways SI
```{r barplot GSEA Fb SI, fig.height=5, fig.width=12}
## select pahtways and filter file
selSI <- c("extracellular matrix organization","regulation of cellular response to growth factor stimulus", "regulation of Wnt signaling pathway", "response to virus")
egoSI_fil <- egoinf%>% filter(egoinf@result$Description %in% selSI)

egoSI_fil_2 <- pairwise_termsim(egoSI_fil)
## make ggbarplot
p <- ggbarplot(egoSI_fil_2@result, x = "Description", y = "Count", fill = "#B45B5C", color= "#B45B5C", sort.val = "asc", orientation = "horizontal")
p
```

### signature umaps SI
#### extracellular matrix organization
```{r plot GSEA extracellular matrix organization}
ego1 <- dplyr::filter(egoinf@result,egoinf@result$Description=="extracellular matrix organization")
g1 <- ego1$geneID
Str <-(g1)
StrSub <- strsplit(Str, "/")
df <- as.data.frame(StrSub)
colnames(df) <- c("gene")

## make a count matrix of signature genes
signGenes <- genes %>% dplyr::filter(gene %in% df$gene)

sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.4))
sceSub$sign2[which(sceSub$sign > 1.4)] <- 1.4
## check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2") + sc +
  theme(legend.position = "none", point_size = 1)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.4))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 1.4)] <- 1.4
## check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.4))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 1.4)] <- 1.4
## check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc

## make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.4))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 1.4)] <- 1.4
## check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

#### regulation of cellular response to growth factor stimulus
```{r plot GSEA regulation of cellular response to growth factor stimulus}
ego1 <- dplyr::filter(egoinf@result,egoinf@result$Description=="regulation of cellular response to growth factor stimulus")
g1 <- ego1$geneID
Str <-(g1)
StrSub <- strsplit(Str, "/")
df <- as.data.frame(StrSub)
colnames(df) <- c("gene")

##make a count matrix of signature genes
signGenes <- genes %>% dplyr::filter(gene %in% df$gene)

sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.2))
sceSub$sign2[which(sceSub$sign > 1.2)] <- 1.2
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2") + sc +
  theme(legend.position = "none", point_size = 1)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.2))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 1.2)] <- 1.2
##check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.2))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 1.2)] <- 1.2
##check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 1.2))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 1.2)] <- 1.2
##check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

### GSEA Fb Naive
#### dotplot
```{r GSEA Fb naive, fig.height=12, fig.width=12}
## genes upregulated in Naive
DEGenesnaive <- dplyr::filter(DEGenesFbcond3, cluster == "naive")
DEGenesnaive <- DEGenesnaive %>%
  mutate(Gene=gsub("^.*\\.", "", gene))  %>%
  mutate(EnsID=gsub("\\..*","", gene))

## GSEA Naive
egonaive <- enrichGO(gene = unique(DEGenesnaive$EnsID),
                          OrgDb = org.Mm.eg.db,
                          keyType = 'ENSEMBL',
                          ont = "BP",
                          pAdjustMethod = "BH",
                          pvalueCutoff = 0.05,
                          qvalueCutoff = 0.05)
egonaive <- setReadable(egonaive, OrgDb = org.Mm.eg.db)
dotplot(egonaive, showCategory=30)
```

#### barplot sel pathways Naive
```{r barplot GSEA Fb naive, fig.height=5, fig.width=12}
## select pahtways and filter file
selnaive <- c("negative regulation of phosphate metabolic process", "response to steroid hormone", "G1/S transition of mitotic cell cycle", "regulation of developmental growth")
egonaive_fil <- egonaive%>% filter(egonaive@result$Description %in% selnaive)

egonaive_fil_2 <- pairwise_termsim(egonaive_fil)
## make ggbarplot
p <- ggbarplot(egonaive_fil_2@result, x = "Description", y = "Count", fill = "#628395", color= "#628395", sort.val = "asc", orientation = "horizontal")
p
```

### signature umaps Naive
#### response to steroid hormone
```{r plot GSEA response to steroid hormone}
ego1 <- dplyr::filter(egonaive@result,egonaive@result$Description =="response to steroid hormone")
g1 <- ego1$geneID
Str <-(g1)
StrSub <- strsplit(Str, "/")
df <- as.data.frame(StrSub)
colnames(df) <- c("gene")

##make a count matrix of signature genes
signGenes <- genes %>% dplyr::filter(gene %in% df$gene)

sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceSub$sign2[which(sceSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2") + sc +
  theme(legend.position = "none", point_size = 1)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

#### G1/S transition of mitotic cell cycle
```{r plot GSEA G1/S transition of mitotic cell cycle}
ego1 <- dplyr::filter(egonaive@result,egonaive@result$Description =="G1/S transition of mitotic cell cycle")
g1 <- ego1$geneID
Str <-(g1)
StrSub <- strsplit(Str, "/")
df <- as.data.frame(StrSub)
colnames(df) <- c("gene")

##make a count matrix of signature genes
signGenes <- genes %>% dplyr::filter(gene %in% df$gene)

sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceSub$sign2[which(sceSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2") + sc +
  theme(legend.position = "none", point_size = 1)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc

##make a count matrix of signature genes naive
sceFbnaiveSub <- sceFbnaive[which(rownames(sceFbnaive) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbnaiveSub@assays@data$logcounts)))/nrow(signGenes)
sceFbnaiveSub$sign <- cntMat
sceFbnaiveSub$sign2 <- sceFbnaiveSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceFbnaiveSub$sign2[which(sceFbnaiveSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceFbnaiveSub$sign)
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbnaiveSub, colour_by = "sign2", point_size = 1) + sc


##make a count matrix of signature genes inf
sceFbinfSub <- sceFbinf[which(rownames(sceFbinf) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbinfSub@assays@data$logcounts)))/nrow(signGenes)
sceFbinfSub$sign <- cntMat
sceFbinfSub$sign2 <- sceFbinfSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceFbinfSub$sign2[which(sceFbinfSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceFbinfSub$sign)
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbinfSub, colour_by = "sign2", point_size = 1) + sc


##make a count matrix of signature genes gran
sceFbgranSub <- sceFbgran[which(rownames(sceFbgran) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceFbgranSub@assays@data$logcounts)))/nrow(signGenes)
sceFbgranSub$sign <- cntMat
sceFbgranSub$sign2 <- sceFbgranSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 0.9))
sceFbgranSub$sign2[which(sceFbgranSub$sign > 0.9)] <- 0.9
##check max and min values
max(sceFbgranSub$sign)
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceFbgranSub, colour_by = "sign2", point_size = 1) + sc
```

#### session info
```{r date and session info}
date()
sessionInfo()
```
