---
title: "GranulomaAll"
author: "A.DeMartin"
date: "2025-08-20"
output:
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#### load packages
```{r load packages, warning=FALSE}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(dplyr)
library(ggplot2)
library(scran)
library(here)
```

### preprocessing
#### load files and merge
```{r load file, eval=FALSE, include=TRUE}
## load objects part1
basedir <- here()
basedir <- paste0(basedir, "/data/dataJuli2022/")
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratN")){
    seuratN <- merge(x = seuratN, y = seuratS)
  }else{
    seuratN <- seuratS
  }
}
remove(seuratS)
table(seuratN$dataset)

## load objects part2
basedir <- here()
basedir <- paste0(basedir, "/data/dataJuli2021/")
fileNamList <- list.files(path = basedir)

for(i in 1:length(fileNamList)){
  seuratS <- readRDS(paste0(basedir, fileNamList[i]))
  if(exists("seuratN")){
    seuratN <- merge(x = seuratN, y = seuratS)
  }else{
    seuratN <- seuratS
  }
}
remove(seuratS)

seuratM <- seuratN
remove(seuratN)
table(seuratM$dataset)

## rerun seurat
seuratM <- NormalizeData (object = seuratM)
seuratM <- FindVariableFeatures(object = seuratM)
seuratM <- ScaleData(object = seuratM, verbose = TRUE)
seuratM <- RunPCA(object=seuratM, npcs = 30, verbose = FALSE)
seuratM <- RunTSNE(object=seuratM, reduction="pca", dims = 1:20)
seuratM <- RunUMAP(object=seuratM, reduction="pca", dims = 1:20)
seuratM <- FindNeighbors(object = seuratM, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratM <- FindClusters(object = seuratM, resolution = res[i], random.seed = 1234)
}

## save seurat object
basedir <- here()
saveRDS(seuratM, file=paste0(basedir, "/data/merged_allgranuloma_wofilter_seurat.rds"))

## remove cluster 21 (0.25res) LEC/Fb doublets
## check counts detected
seuratLEC <- subset(seurat, RNA_snn_res.0.25 =="21")
sce <- as.SingleCellExperiment(seuratLEC)
plotQC2 <- function(sce, feature){
  p <- gghistogram(data = as.data.frame(sce@colData),
              x=feature,
              bins=70,
              #title=unique(sce$dataset),
              fill = "slategray",
              legend = "right", color= "black")
  return(p)
}
plotQC2(sce, "total")
plotQC2(sce, "detected")

## remove 
table(seurat$RNA_snn_res.0.25)
seurat1 <- subset(seurat, RNA_snn_res.0.25 == "21", invert = TRUE)
table(seurat1$RNA_snn_res.0.25)

## save seurat object
saveRDS(seurat, file=paste0(basedir, "/data/merged_allgranuloma_wofilter_seurat_wocl21.rds"))
```

#### add metadata
```{r add metadata, eval=FALSE, include=TRUE}
## cluster_name
seurat$cluster_name <- "cluster_name"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "0" )] <- "PdgfraloFb1"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "1" )] <- "Mph1"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "2" )] <- "CD8Tcell1"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "3" )] <- "Neut"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "4" )] <- "IgABcell1"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "5" )] <- "LEC"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "6" )] <- "Telocytes"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "7" )] <- "DC"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "8" )] <- "CD4Tcell/ILC2"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "9" )] <- "Mph2"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "10" )] <- "NKcell"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "11" )] <- "CD8Tcell2"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "12" )] <- "Mph3"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "13" )] <- "Trophocytes"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "14" )] <- "PdgfrahiFb"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "15" )] <- "IgABcell2"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "16" )] <- "Myocytes"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "17" )] <- "PdgfraloFb2"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "18" )] <- "IgDBcell"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "19" )] <- "Glial"
seurat$cluster_name[which(seurat$RNA_snn_res.0.25 %in% "20" )] <- "ICC"
table(seurat$cluster_name)
table(seurat$RNA_snn_res.0.25)

## order
Idents(seurat) <- seurat$cluster_name
seurat$cluster_name <- factor(seurat$cluster_name, levels=c("CD8Tcell1","CD8Tcell2","NKcell", "CD4Tcell/ILC2","Neut", "Mph1","Mph2", "Mph3","DC","IgABcell1","IgABcell2","IgDBcell","PdgfraloFb1","PdgfraloFb2", "Trophocytes", "Telocytes", "Myocytes", "PdgfrahiFb", "Glial", "LEC", "ICC"))
Idents(seurat) <- seurat$cluster_name
table(seurat$cluster_name)

## add slot cluster plus cond2
seurat$cluster_plus_cond2 <- paste0(seurat$seurat_clusters, '_', seurat$cond2)

## cond2
WT <- c("o289411_1-1_20220629_Mu_granuloma_Cxcl13Cre_CD45_HpD4_v3","o289411_2-2_20220629_Mu_granuloma_Cxcl13Cre_FSC_HpD4_v3","9_20210624_Mu_granuloma_Cxcl13Cre_EYFP_HpD4_v3")
cko <- c("o289411_3-3_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_CD45_HpD4_v3","o289411_4-4_20220629_Mu_granuloma_Cxcl13Cre_IL33fl_FSC_HpD4_v3","10_20210624_Mu_granuloma_Cxcl13Cre_IL33flf_EYFP_HpD4_v3")

seurat$cond2 <- "cond2"
seurat$cond2[which(seurat$dataset %in% WT)] <- "WT"
seurat$cond2[which(seurat$dataset %in% cko)] <- "cko"
table(seurat$cond2)

## order
Idents(seurat) <- seurat$cond2
seurat$cond2 <- factor(seurat$cond2, levels=c("WT", "cko"))
Idents(seurat) <- seurat$cond2
table(seurat$cond2)

## save seurat object
saveRDS(seurat, file=paste0(basedir, "/data/merged_allgranuloma_wofilter_seurat_wocl21.rds"))
```

### end preprocessing
#### load merged file granuloma
```{r load merged file granuloma}
basedir <- here()
fileNam <- paste0(basedir, "/data/merged_allgranuloma_wofilter_seurat_wocl21.rds")
seurat <- readRDS(fileNam)
```

#### counts
```{r}
table(seurat$orig.ident)
table(seurat$cond2)
```

#### set color vectors
```{r set color vectors}
## col clustername
colclustername <- c("#355C7D","#B1746FFF","#202547","#B09C85", "#4e5a4c","#53354A","#2A363B","#8491B4FF","#00A087FF","#DC9989","#84ad83", "#628395","#D33B44","#779d8d","#727077","#868686FF","#F8B195","#FF847C","#725663FF","#904D39","#91D1C2")
names(colclustername) <- c("PdgfraloFb1","Mph1", "CD8Tcell1","Neut", "IgABcell1", "LEC", "Telocytes", "DC", "CD4Tcell/ILC2", "Mph2",  "NKcell", "CD8Tcell2", "Mph3", "Trophocytes", "PdgfrahiFb", "IgABcell2", "Myocytes", "PdgfraloFb2", "IgDBcell", "Glial", "ICC")

## col cond2
colcond2 <- c("#202547","#BE3144")
names(colcond2) <- c("WT", "cko")
```

### umaps
#### clustername
```{r plot umaps clustername}
Idents(seurat) <- seurat$cluster_name
DimPlot(seurat, reduction = "umap", cols = colclustername)
DimPlot(seurat, reduction = "umap", cols = colclustername) + theme(legend.position = "none")
```

#### condition
```{r plot umaps condition}
Idents(seurat) <- seurat$cond2
DimPlot(seurat, reduction = "umap", cols = colcond2, split.by = "cond2")
DimPlot(seurat, reduction = "umap", cols = colcond2, order = "WT")
DimPlot(seurat, reduction = "umap", cols = colcond2, order="WT") + theme(legend.position = "none")
```

#### calculate marker genes
```{r calculate marker genes, eval=FALSE, include=TRUE}
## cluster marker genes
Idents(seurat) <- seurat$cluster_name
levels(seurat)

markerGenes <- FindAllMarkers(seurat, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01) 

write.table(markerGenes, 
            file= paste0(basedir, "/data/markergenes_cluster_name"),
            sep="\t",
            quote=F,
            row.names=F,
            col.names=T)
```

### dotplot cluster marker
```{r dotplot, fig.height=11, fig.width=10}
genes <- data.frame(gene=rownames(seurat)) %>% 
  mutate(geneID=gsub("^.*\\.", "", gene))

selGenes <- data.frame(geneID=c("Cpe","Ano1","Pecam1","Prox1","Lyve1","Sostdc1","Plp1","Ly6a","Acta2","Cd81","Cd34","Ncam1","Pdgfra","Pdpn", "Cd79a", "Cd79b", "Cd19", "Ighm", "Ighd", "Igha","Itgax","H2-Eb1","H2-Ab1","Cd74","Siglech","Ly6c1","Adgre1","Cx3cr1","Lyz2","Itgam","Msr1", "Arg1", "Ly6g","S100a8", "S100a9", "Il1b", "Rora", "Gata3", "Icos", "Cd4", "Klre1", "Klrk1", "Nkg7", "Gzmb","Cd8b1","Cd8a", "Trbc2", "Cd3d", "Trbc1")) %>% left_join(., genes, by="geneID")

DotPlot(seurat, features = selGenes, group.by= "cluster_name") + RotatedAxis() + scale_color_viridis(option="E") + coord_flip()
```

### subset CD45 negative clusters
```{r subset CD45neg}
seuratCD45neg <- subset(seurat, cluster_name %in% c("PdgfraloFb1","PdgfraloFb2", "Trophocytes", "Telocytes", "Myocytes", "PdgfrahiFb", "Glial", "LEC", "ICC"))
```

#### counts
```{r}
table(seuratCD45neg$cluster_name)
table(seuratCD45neg$orig.ident)
table(seuratCD45neg$cond2)
```

#### average Heatmap top5 cluster marker CD45 negative (stromal)
```{r avg heatmap, fig.height=10, fig.width=8}
Idents(seuratCD45neg) <- seuratCD45neg$cluster_name
levels(seuratCD45neg)

avgHeatmap <- function(seuratCD45neg, selGenes, colVecIdent, colVecCond=NULL,
                       ordVec=NULL, gapVecR=NULL, gapVecC=NULL,cc=FALSE,
                       cr=FALSE, condCol=FALSE){
  
  selGenes <- selGenes$gene
  
  ## assay data
  clusterAssigned <- as.data.frame(Idents(seuratCD45neg)) %>%
  dplyr::mutate(cell=rownames(.))
  colnames(clusterAssigned)[1] <- "ident"
  seuratDat <- GetAssayData(seuratCD45neg)
  
  ## genes of interest
  genes <- data.frame(gene=rownames(seuratCD45neg)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) %>% filter(geneID %in% selGenes)

  ## matrix with averaged cnts per ident
  logNormExpres <- as.data.frame(t(as.matrix(
    seuratDat[which(rownames(seuratDat) %in% genes$gene),])))
  logNormExpres <- logNormExpres %>% dplyr::mutate(cell=rownames(.)) %>%
    dplyr::left_join(.,clusterAssigned, by=c("cell")) %>%
    dplyr::select(-cell) %>% dplyr::group_by(ident) %>%
    dplyr::summarise_all(mean)
  logNormExpresMa <- logNormExpres %>% dplyr::select(-ident) %>% as.matrix()
  rownames(logNormExpresMa) <- logNormExpres$ident
  logNormExpresMa <- t(logNormExpresMa)
  rownames(logNormExpresMa) <- gsub("^.*?\\.","",rownames(logNormExpresMa))
  
  ## remove genes if they are all the same in all groups
  ind <- apply(logNormExpresMa, 1, sd) == 0
  logNormExpresMa <- logNormExpresMa[!ind,]
  genes <- genes[!ind,]

  ## color columns according to cluster
  annotation_col <- as.data.frame(gsub("(^.*?_)","",
                                       colnames(logNormExpresMa)))%>%
    dplyr::mutate(celltype=gsub("(_.*$)","",colnames(logNormExpresMa)))
  colnames(annotation_col)[1] <- "col1"
  annotation_col <- annotation_col %>%
    dplyr::mutate(cond = gsub("(^[0-9]_?)","",col1)) %>%
    dplyr::select(cond, celltype)
  rownames(annotation_col) <- colnames(logNormExpresMa) 

  ann_colors = list(
      cond = colVecCond,
      celltype=colclustername)
  if(is.null(ann_colors$cond)){
    annotation_col$cond <- NULL
  }
  
  ## adjust order
  logNormExpresMa <- logNormExpresMa[selGenes,]
  if(is.null(ordVec)){
    ordVec <- levels(seuratCD45neg)
  }
  logNormExpresMa <- logNormExpresMa[,ordVec]

  ## scaled row-wise
  pheatmap(logNormExpresMa, scale="row" ,treeheight_row = 0, cluster_rows = cr, 
         cluster_cols = cc,
         color = colorRampPalette(c("#2166AC", "#F7F7F7", "#B2182B"))(50),
         annotation_col = annotation_col, cellwidth=15, cellheight=10,
         annotation_colors = ann_colors, gaps_row = gapVecR, gaps_col = gapVecC)
}

genesPlot <- data.frame(gene=c("Sfrp1", "Clec3b", "Fbln1", "Adamdec1", "C4b", "P2rx1", "Itga8", "Fgfr2", "Cxcl14", "Tmem178", "Cpxm1", "Ccdc80", "Cd34", "Ogn", "Lgi2", "Bmp5", "Tagln", "Tmem158", "Col4a5", "Smoc2", "Myh11", "Actg2", "Acta2", "Myl9", "Tpm2", "Ptn", "Rgs5", "Dpt", "Lum", "Timp1", "Plp1", "Lgi4", "Cryab", "Sostdc1", "Postn", "Mmrn1", "Flt4", "Egfl7", "Lyve1", "Aqp1", "Cpe", "Ano1", "Pde3a", "Pcp4l1", "Grem2"))

levels(seuratCD45neg)
colVec <- colclustername

avgHeatmap(seuratCD45neg, selGenes = genesPlot, colVecIdent = colVec)
```

### abundances
#### CD45 negative (stromal)
```{r abundance CD45neg}
## make count list
datList <- NULL
for(con in unique(seuratCD45neg$cond2)){
  seuratSub <- subset(seuratCD45neg, cond2==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$cluster_name)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

## order x
ordX <-  c("WT", "cko") 

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclustername) + scale_x_discrete(limits=ordX)
```

#### CD45 positive (immmune cells)
```{r}
## subset CD45+ cells
seuratCD45pos <- subset(seurat, cluster_name %in% c("CD8Tcell1","CD8Tcell2","NKcell","CD4Tcell/ILC2","Neut","Mph1","Mph2","Mph3","DC","IgABcell1","IgABcell2","IgDBcell"))
table(seuratCD45pos$cluster_name)
table(seuratCD45pos$orig.ident)
```

#### counts
```{r}
table(seuratCD45pos$cluster_name)
table(seuratCD45pos$orig.ident)
table(seuratCD45pos$cond2)
```

```{r relative abundance CD45pos}
## make count list
datList <- NULL
for(con in unique(seuratCD45pos$cond2)){
  seuratSub <- subset(seuratCD45pos, cond2==con)
  print(dim(seuratSub))
  dat_con <- as.data.frame(table(seuratSub$cluster_name)) %>%
  mutate(percent=Freq/ncol(seuratSub)) %>% mutate(cond=con)
  datList[[con]] <- dat_con
}
dat_all <- do.call("rbind", datList)

## order x
ordX <-  c("WT", "cko") 

## plot abundance
ggbarplot(dat_all, x= "cond", y= "percent", fill = "Var1", legend = "right", legend.titel = "cluster", ylab = "frequency", palette = colclustername) + scale_x_discrete(limits=ordX)
```

### top DE genes
#### assign cell types
```{r assign celltypes}
seurat$celltype <- "celltype"
seurat$celltype[which(seurat$cluster_name %in% c("CD8Tcell1", "CD8Tcell2"))] <- "CD8Tcells"
seurat$celltype[which(seurat$cluster_name %in% c("NKcell"))] <- "NKcells"
seurat$celltype[which(seurat$cluster_name %in% c("CD4Tcell/ILC2"))] <- "CD4Tcell/ILC2"
seurat$celltype[which(seurat$cluster_name %in% c("Neut"))] <- "Neut"
seurat$celltype[which(seurat$cluster_name %in% c("Mph1", "Mph2", "Mph3"))] <- "Macrophages"
seurat$celltype[which(seurat$cluster_name %in% c("DC"))] <- "DC"
seurat$celltype[which(seurat$cluster_name %in% c("IgABcell1","IgABcell2","IgDBcell"))] <- "Bcells"
seurat$celltype[which(seurat$cluster_name %in% c("PdgfraloFb1","PdgfraloFb2", "Trophocytes", "Telocytes", "Myocytes", "PdgfrahiFb"))] <- "Fb"
seurat$celltype[which(seurat$cluster_name %in% c("Glial"))] <- "Glial"
seurat$celltype[which(seurat$cluster_name %in% c("LEC"))] <- "LEC"
seurat$celltype[which(seurat$cluster_name %in% c("ICC"))] <- "ICC"
table(seurat$celltype)

## order
Idents(seurat) <- seurat$celltype
seurat$celltype <- factor(seurat$celltype, levels=c("CD8Tcells","NKcells", "CD4Tcell/ILC2","Neut","Macrophages","DC","Bcells", "Fb", "Glial", "LEC", "ICC"))
Idents(seurat) <- seurat$celltype
table(seurat$celltype)
```
#### set color vector celltypes
```{r set color vectors celltype}
colcelltype <- c("#202547","#84ad83","#00A087FF","#B09C85","#B1746FFF","#8491B4FF", "#4e5a4c","#355C7D","#904D39","#725663FF","#91D1C2")
names(colcelltype) <- c("CD8Tcells","NKcells", "CD4Tcell/ILC2","Neut","Macrophages","DC","Bcells", "Fb", "Glial", "LEC", "ICC")
```

### umap
#### celltype
```{r celltype umap}
Idents(seurat) <- seurat$celltype
DimPlot(seurat, reduction = "umap", cols = colcelltype)
DimPlot(seurat, reduction = "umap", cols = colcelltype) + theme(legend.position = "none")
```

#### calculate DE genes for each immune celltype according to WT/cko
```{r DE genes celltypes}
Idents(seurat) <- seurat$cond2

## top 100 DE genes CD8Tcells
seuratCD8Tcells <- subset(seurat, celltype == "CD8Tcells")
table(seuratCD8Tcells$cond2)
levels(seuratCD8Tcells)

DEGenesCD8TcellsWTcko <- FindAllMarkers(seuratCD8Tcells, logfc.threshold = 0.1) %>% 
  dplyr::filter(p_val_adj < 0.1) %>% mutate(celltype = "CD8Tcells")
DEGenesCD8TcellsWTcko100 <- DEGenesCD8TcellsWTcko %>% top_n(100, avg_log2FC) 

## top 100 DE genes NKcells
seuratNKcells <- subset(seurat, celltype == "NKcells")
table(seuratNKcells$cond2)
levels(seuratNKcells)

DEGenesNKcellsWTcko <- FindAllMarkers(seuratNKcells,logfc.threshold = 0.1) %>% 
  dplyr::filter(p_val_adj < 0.1) %>% mutate(celltype = "NKcells")
DEGenesNKcellsWTcko100 <- DEGenesNKcellsWTcko %>% top_n(100, avg_log2FC) 

## top 100 DE genes CD4Tcell/ILC2
seuratCD4TcellILC2 <- subset(seurat, celltype == "CD4Tcell/ILC2")
table(seuratCD4TcellILC2$cond2)
levels(seuratCD4TcellILC2)

DEGenesCD4TcellILC2WTcko <- FindAllMarkers(seuratCD4TcellILC2,logfc.threshold = 0.1) %>% 
  dplyr::filter(p_val_adj < 0.1) %>% mutate(celltype = "CD4Tcell/ILC2")
DEGenesCD4TcellILC2WTcko100 <- DEGenesCD4TcellILC2WTcko %>% top_n(100, avg_log2FC)

## top 100 DE genes Neut
seuratNeut <- subset(seurat, celltype == "Neut")
table(seuratNeut$cond2)
levels(seuratNeut)

DEGenesNeutWTcko <- FindAllMarkers(seuratNeut,logfc.threshold = 0.1) %>% 
  dplyr::filter(p_val_adj < 0.1) %>% mutate(celltype = "Neut")
DEGenesNeutWTcko100 <- DEGenesNeutWTcko %>% top_n(100, avg_log2FC) 

## top 100 DE genes Macrophages
seuratMacrophages <- subset(seurat, celltype == "Macrophages")
table(seuratMacrophages$cond2)
levels(seuratMacrophages)

DEGenesMacrophagesWTcko <- FindAllMarkers(seuratMacrophages,logfc.threshold = 0.1) %>% 
  dplyr::filter(p_val_adj < 0.1) %>% mutate(celltype = "Macrophages")
DEGenesMacrophagesWTcko100 <- DEGenesMacrophagesWTcko %>% top_n(100, avg_log2FC) 

## top 100 DE genes DC
seuratDC <- subset(seurat, celltype == "DC")
table(seuratDC$cond2)
levels(seuratDC)

DEGenesDCWTcko <- FindAllMarkers(seuratDC, logfc.threshold = 0.1) %>% 
  dplyr::filter(p_val_adj < 0.1) %>% mutate(celltype = "DC")
DEGenesDCWTcko100 <- DEGenesDCWTcko %>% top_n(100, avg_log2FC) 

## top 100 DE genes Bcells
seuratBcells <- subset(seurat, celltype == "Bcells")
table(seuratBcells$cond2)
levels(seuratBcells)

DEGenesBcellsWTcko <- FindAllMarkers(seuratBcells,logfc.threshold = 0.1) %>% 
  dplyr::filter(p_val_adj < 0.1) %>% mutate(celltype = "Bcells")
DEGenesBcellsWTcko100 <- DEGenesBcellsWTcko %>% top_n(100, avg_log2FC) 

## joint all
DEGenesAll100WTcko <- full_join(DEGenesCD8TcellsWTcko100, DEGenesNKcellsWTcko100)
DEGenesAll100WTcko <- full_join(DEGenesAll100WTcko, DEGenesCD4TcellILC2WTcko100)
DEGenesAll100WTcko <- full_join(DEGenesAll100WTcko, DEGenesNeutWTcko100)
DEGenesAll100WTcko <- full_join(DEGenesAll100WTcko, DEGenesMacrophagesWTcko100)
DEGenesAll100WTcko <- full_join(DEGenesAll100WTcko, DEGenesDCWTcko100)
DEGenesAll100WTcko <- full_join(DEGenesAll100WTcko, DEGenesBcellsWTcko100)
```

#### distribution of log2FC of overall top 100 DE genes
```{r boxplot top100 DE genes overall}
ggdensity(DEGenesAll100WTcko, x = "avg_log2FC", add= "median", rug = TRUE, color = "celltype", fill = "celltype", palette = colcelltype)

ggviolin(DEGenesAll100WTcko, x = "celltype", y = "avg_log2FC", fill = "celltype", palette = colcelltype, add = "median_iqr")

ggboxplot(DEGenesAll100WTcko, x = "celltype", y = "avg_log2FC", color = "celltype", palette = colcelltype ,add = "jitter")
```

### violin Il1rl1
```{r violin}
Idents(seurat) <- seurat$cluster_name
VlnPlot(object=seurat, features = "ENSMUSG00000026069.Il1rl1", cols = colcond2, split.by = "cond2", pt.size = 1)
VlnPlot(object=seurat, features = "ENSMUSG00000026069.Il1rl1", cols = colcond2, split.by = "cond2", pt.size = 2)
```

#### session info
```{r date and session info}
date()
sessionInfo()
```
